R Notebook

### 2018-05-05 analysis/final_analysis.Rmd
Exclude buggy trials; look at ref exps

```{r}
exp1_messages <- read_csv('../data/final/experiment1/chatMessage/messagesFromMongo.csv') %>%
  mutate(critical = ifelse(critical == 'True', 1, 0)) %>%
  rename(messageTime = serverTime) %>%
  filter(sender == 'director') %>%
  group_by(gameid, trialNum, instructionNum, attemptNum, critical, condition, objectSet, targetObject) %>%
  summarize(text = paste0(contents, collapse = ' '),
            typingRT = sum(typingRT)) %>%   # Concatenate multiple utterance on same trial
  mutate(numRawWords = str_count(text, "\\S+"))

## Have to fix attemptNum bug (got recorded incorrectly)
exp1_drops <- read_csv('../data/final/experiment1/drop/dropsFromMongo.csv') %>%
  mutate(critical = ifelse(critical == 'True', 1, 0),
         correct = ifelse(correct == 'correct', 1, 0),
         instructionNum = ifelse(correct == 0, instructionNum + 1, instructionNum)) %>%
  group_by(gameid, instructionNum, trialNum) %>%
  mutate(attemptNum = ifelse(attemptNum == 0, max(attemptNum), attemptNum - 1)) %>%
  rename(dropTime = serverTime) %>%
  select(-X1, -eventType)
```

@@

exp1_badGames <- c(incompleteIDs, nonNative, confused, nonCriticalMistakes)

exp1_d <- exp1_messages %>%
  left_join(exp1_drops) %>% #, by = c('gameid', 'trialNum', 'instructionNum', 'trialType', 
                      #         'targetObject', 'critical', 'condition', 'attemptNum')) %>%
  filter(!(gameid %in% exp1_badGames))
```

@@

@@

critTrials %>%
  group_by(gameid, condition) %>%
  summarize(pctCriticalError = 1 - mean(correct)) %>%
  ggplot(aes(x = pctCriticalError, fill = condition)) +
    #geom_density(aes(y = ..density../sum(..density..)), alpha = 0.5, adjust = 1.25)
    geom_histogram(aes(y = ..count../sum(..count..)), alpha = 0.5, position = 'dodge', binwidth = 0.125)  +
    facet_grid(condition ~ .) +
    theme_few()

ggsave('~/Downloads/criticalErrors.pdf')
```

@@

ggsave('~/Downloads/itemHeterogeneity.pdf')
```

@@

dodge <- position_dodge(width=0.9)
d.onDistractor %>%
  group_by(condition, trialType) %>%
  tidyboot_mean(column = hovered) %>%
  ggplot(aes(x = condition, y=empirical_stat, fill = trialType)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) +
    theme_few() +
    ylab('% trials hovering over distractor') +
    theme(aspect.ratio = 0.5)

ggsave('~/Downloads/hoverOverDistractor.pdf')
```

@@

@@

# Some speakers relied on location in grid (TODO: WHAT ABOUT 'BELOW' PEOPLE?)
location_abusers <- c('8219-bb7861c1-43f4-480c-906b-8441c583c0ab', 
                      '7726-3d4a266f-e56d-4afb-9a79-b40deb690a76', 
                      '5684-f09de856-4d64-423c-b342-b30e7325eb0e', 
                      '7600-346d56b5-bd76-406f-befb-6b47e07d5916')
# Some speakers played this weird taboo game where they gave riddles 
tabooers <- c('2929-d218f724-b45e-416b-af44-5cab5658bad9', #(opposite of white... primary color...)
              '7600-346d56b5-bd76-406f-befb-6b47e07d5916', # MERRY GO ROUND
              '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???
badGames <- c(incompleteIDs, pureLocation, confused)

@@

ggplot(basic.d, aes(x = far, y = close)) +
    geom_point() +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(1,12) +
    xlim(1,12) +
    theme(aspect.ratio = 1)
```

@@

```{r}
ggplot(basic.d, aes(x = ratio)) +
    geom_histogram(binwidth = 0.25) +
    theme_bw() +
    geom_vline(xintercept = 0) +
    # ylim(1,12) +
    xlim(-4,4) +
    theme(aspect.ratio = 1)

t.test(basic.d$ratio)
```

@@

d %>%
  group_by(context, hidden) %>%
  tidyboot_mean(column = numRawWords) %>%
  ggplot(aes(x = context, y = empirical_stat, fill = hidden)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
    theme_few() +
    ylab('mean # words') +
    theme(aspect.ratio = .5) +
    scale_fill_colorblind()

@@

ggplot(hidden.d, aes(x = no, y = yes)) +
    geom_point() +
    facet_wrap(~ context) +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(1,15) +
    ylab('# words (with occlusions)') +
    xlab('# words (without occlusions)') +
    ggtitle('Within-pair comparisons of reference w/ and w/out occlusion') +
    xlim(1,15) +
    theme(aspect.ratio = 1)

@@

```{r}
ggplot(hidden.d %>% filter(!(gameid %in% all_the_things_ers)) %>% group_by(context) %>% mutate(m = mean(diff)), aes(x = diff)) +
    geom_histogram(binwidth = 0.33) +
    theme_bw() +
    geom_vline(xintercept = 0, linetype = 2) +
    geom_vline(aes(xintercept = m), color = 'red') +
    facet_wrap(~ context) +
    # ylim(1,12) +
    xlim(-6,6) +
    theme(aspect.ratio = 1)

@@

```{r}
hidden.d %>%
  group_by(context) %>%
  summarize(m = mean(diff), se = sd(diff)/sqrt(length(diff))) %>%
  ggplot(aes(x = context, y = m)) +
    geom_bar(stat = 'identity') +
    geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0) +
    theme_few() +
    ylim(-3, 3) +
    ylab('# words (occlusions - no occlusions)')

@@

ggplot(d.booted, aes(x = context, y = empirical_stat, fill = hidden)) + 
  geom_bar(stat = 'identity', position = dodge) +
  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
  facet_wrap(~ feature) + 
  theme_few() +
  theme(aspect.ratio = 1) + 
  ylab('% utterances with feature') +
  scale_fill_colorblind()

>>>>

```{r}
raw <- read_csv('../data/final/experiment1/chatMessage/messagesFromMongo.csv')
exp1_messages <- read_csv('../data/final/experiment1/chatMessage/messagesFromMongo.csv') %>%
  mutate(critical = ifelse(critical == 'True', 1, 0)) %>%
  rename(messageTime = serverTime) %>%
  filter(sender == 'director') %>%
  group_by(gameid, trialNum, instructionNum, attemptNum, critical, condition, objectSet, targetObject) %>%
  summarize(text = paste0(contents, collapse = ' '),
            typingRT = sum(typingRT)) %>%   # Concatenate multiple utterance on same trial
  mutate(numRawWords = str_count(text, "\\S+"))

## Have to post-hoc attemptNum and instructionNum bugs (got reset for next trial prematurely)
## Also, remove trials where multi-drop bug happened (i.e. where clicking multiple times triggered multiple instruction resetting)
exp1_drops <- read_csv('../data/final/experiment1/drop/dropsFromMongo.csv') %>%
  mutate(critical = ifelse(critical == 'True', 1, 0),
         correct = ifelse(correct == 'correct', 1, 0),
         instructionNum = ifelse(correct == 0, instructionNum + 1, instructionNum)) %>%
  group_by(gameid, targetObject) %>%
  filter(length(unique(instructionNum)) < 2) %>%
  group_by(gameid, instructionNum, trialNum) %>%
  mutate(attemptNum = ifelse(attemptNum == 0, max(attemptNum), attemptNum - 1)) %>%
  rename(dropTime = serverTime) %>%
  select(-X1, -eventType)
```

@@

buggyTrials <- read_csv('../data/final/experiment1/drop/dropsFromMongo.csv') %>%
  mutate(critical = ifelse(critical == 'True', 1, 0),
       correct = ifelse(correct == 'correct', 1, 0),
       instructionNum = ifelse(correct == 0, instructionNum + 1, instructionNum)) %>%
  group_by(gameid, targetObject) %>%
  mutate(numInstructions = length(unique(instructionNum))) %>%
  filter(numInstructions > 1)

taboo <- c('4241-cda4a373-ffd7-41e3-ae60-82b9d377a455')
onlyLocation <- c('3222-f68d2df5-8977-4ff2-9ee9-c60f8ddb61f1', '9457-4189cc9b-a82b-4db9-828d-bf90d1987aee')
exp1_badGames <- c(incompleteIDs, nonNative, confused, nonCriticalMistakes, taboo, onlyLocation, unique(buggyTrials$gameid))

exp1_d <- exp1_drops %>% 
  left_join(exp1_messages) %>% #, by = c('gameid', 'trialNum', 'instructionNum', 'trialType', 
                      #         'targetObject', 'critical', 'condition', 'attemptNum')) %>%
  filter(!(gameid %in% exp1_badGames))
```

@@

And slightly uneven numbers of 'experimental' trials in each item...

```{r}
exp1_d %>% 
  group_by(gameid, objectSet, trialType, condition) %>% 
  tally() %>% 
  group_by(condition, objectSet, trialType) %>% 
  summarize(numInCondition = length(n))
```

@@

critTrials %>%
  group_by(gameid, condition) %>%
  summarize(pctCriticalError = 4 - sum(correct)) %>%
  ggplot(aes(x = pctCriticalError, fill = condition)) +
    geom_histogram(alpha = 0.5, position = 'dodge', binwidth = 0.5)  +
    facet_grid(condition ~ .) +
    theme_few() +
    ylab('# pairs making k critical errors') +
    xlab('# critical errors') + 
    theme(aspect.ratio = 1, legend.position="none") +
    scale_fill_colorblind()+
    ggtitle('critical errors')

ggsave('../writing/journal_manuscript/figures/criticalErrors.pdf')
```

```{r}
critTrials %>%
  group_by(condition) %>%
  mutate(error = 1 - correct) %>%
  tidyboot_mean(column = error) %>%
  ggplot(aes(x = condition, y =empirical_stat, fill = condition)) +
    geom_bar(stat = 'identity', alpha = .5) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
    ylab('% errors') +
    scale_fill_colorblind()+
    theme_few() +
    theme(aspect.ratio = 1, legend.position="none")

ggsave('../writing/journal_manuscript/figures/errorMeans.pdf', height = 3, width = 3)
```

@@

ggsave('../writing/journal_manuscript/figures//itemHeterogeneity.pdf')
```

@@

dodge <- position_dodge(width=0.9)
d.onDistractor %>%
  group_by(condition, trialType) %>%
  tidyboot_mean(column = hovered) %>%
  ggplot(aes(x = condition, y=empirical_stat, fill = trialType)) +
    geom_bar(stat = 'identity', position = dodge, alpha =.5) +
    geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0, position = dodge) +
    theme_few() +
    ylab('% trials hovering over distractor') +
    scale_fill_colorblind()+
    theme(aspect.ratio = 0.5)

ggsave('../writing/journal_manuscript/figures//hoverOverDistractor.pdf', height = 3, width = 6)
```

@@

## Prep informativity norming stims

```{r}
refExps <- exp1_d %>% 
  filter(condition == 'unscripted') %>% 
  filter(critical == 1) %>% 
  filter(attemptNum == 0) %>% 
  ungroup() %>%
  select(gameid, objectSet, text)

write_csv(refExps, './refExpressions_raw.csv')
```

For 6 of the 8 items, unscripted speakers had significantly more words in their referring expressions... (n.s. for the other 2, but idiosyncratic)

```{r}
refs <- read_csv('../data/final/experiment1/refExpressions.csv')
scripted <- data.frame(objectSet = c(1,2,3,4,5,6,7,8), 
                       m = c('glasses', 'bottom block', 'tape', 'large measuring cup', 'brush', 'eraser', 'small candle', 'mouse'),
                       numRawWords = c(1,2,1,3,1,1,2,1))
refs %>% 
  mutate(numRawWords = str_count(text, "\\S+")) %>%
  group_by(objectSet) %>%
  tidyboot_mean(numRawWords) %>%
  left_join(scripted)
```

## TO-DO

1. Subject-level analyses of informativity: could be a mixture of 'lazy' egocentric speakers and 'attentive' audience-sensitive speakers? 
2. Sometimes the additional information is in movement instruction (i.e. "one square left toward a curtain" or "one square up below the comb")
3. Change in informativity over time (are people already being informative on first round? are people 'learning' to be more informative over course of expt? or getting less informative because tired of task?)
4. Just length of referring expression in words?

@@

# Some speakers relied on location in grid (TODO: WHAT ABOUT 'BELOW' PEOPLE?)
location_abusers <- c('8219-bb7861c1-43f4-480c-906b-8441c583c0ab', 
                      '7726-3d4a266f-e56d-4afb-9a79-b40deb690a76', 
                      '5684-f09de856-4d64-423c-b342-b30e7325eb0e', 
                      '7600-346d56b5-bd76-406f-befb-6b47e07d5916')
# Some speakers played this weird taboo game where they gave riddles 
tabooers <- c('2929-d218f724-b45e-416b-af44-5cab5658bad9', #(opposite of white... primary color...)
              '7600-346d56b5-bd76-406f-befb-6b47e07d5916', # MERRY GO ROUND
              '3462-7e95c955-4087-4eb9-bbc8-05338f9a5e4b') # CAN YOU SEE???
badGames <- c(incompleteIDs, location_abusers, confused)

@@

ggplot(basic.d, aes(x = far, y = close)) +
    geom_point() +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(1,12) +
    xlim(1,12) +
    theme(aspect.ratio = 1, text = element_text(size=20))
```

@@

```{r}
tidyboot_mean(basic.d %>%ungroup(), column = diff) %>%
  ggplot(aes(x = 'close - far', y = empirical_stat)) +
      geom_bar(stat ='identity', width = 0.25) +
      geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), width = 0) +
      theme_few()+
      geom_hline(yintercept = 0) +
      #geom_vline(xintercept = 0) +
      ylim(-1,1) +
      # xlim(-4,4) +
      ylab('# words used') +
    theme(aspect.ratio = 1, text = element_text(size=20))

t.test(basic.d$diff)
```

@@

d %>%
  group_by(context, hidden) %>%
  tidyboot_mean(column = numRawWords) %>%
  ggplot(aes(x = context, y = empirical_stat, fill = hidden)) +
    geom_bar(stat = 'identity', position = dodge) +
    geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = dodge, width = 0) +
    theme_few() +
    ylab('mean # words') +
    theme(aspect.ratio = 1, text = element_text(size=20)) +
    scale_fill_colorblind()

@@

ggplot(hidden.d, aes(x = no, y = yes)) +
    geom_point() +
    facet_wrap(~ context) +
    theme_bw() +
    geom_abline(intercept = 0, slope = 1) +
    ylim(1,15) +
    ylab('# words (with occlusions)') +
    xlab('# words (without occlusions)') +
    #ggtitle('Within-pair comparisons of reference w/ and w/out occlusion') +
    xlim(1,15) +
    theme(aspect.ratio = 1, text = element_text(size=20))

@@

```{r}
hidden.d %>% 
  #filter(!(gameid %in% all_the_things_ers)) %>% 
  group_by(context) %>% 
  mutate(m = mean(diff)) %>%
  ggplot(aes(x = diff)) +
      geom_histogram(binwidth = 0.33) +
      theme_bw() +
      geom_vline(xintercept = 0, linetype = 2) +
      geom_vline(aes(xintercept = m), color = 'red') +
      facet_wrap(~ context) +
      # ylim(1,12) +
      xlim(-6,6) +
      theme(aspect.ratio = 1)

@@

```{r}
hidden.d %>%
  group_by(context) %>%
  summarize(m = mean(diff), se = sd(diff)/sqrt(length(diff))) %>%
  ggplot(aes(x = context, y = m)) +
    geom_bar(stat = 'identity', width = 0.25) +
    geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0) +
    theme_few() +
    geom_hline(yintercept = 0) +
    ylim(-3, 3) +
    theme(aspect.ratio = 1, text = element_text(size=20))  +
    ylab('# words (occlusions - no occlusions)')

@@

ggplot(d.booted, aes(x = context, y = empirical_stat, fill = hidden)) + 
  geom_bar(stat = 'identity', position = dodge) +
  geom_errorbar(aes(ymax = ci_upper, ymin = ci_lower), position = dodge, width = 0) +
  facet_wrap(~ feature) + 
  theme_few() +
  theme(aspect.ratio = 1, text = element_text(size=20))  +
  ylab('% utterances with feature') +
  scale_fill_colorblind()
